---
title: "Probability: Basic Definitions and Rules"
author: "Ognjen Raketic"
format: pdf
editor: visual
jupyter: python3
---

## Probability: Basic Definitions and Rules

-   In the last lecture we already discussed in the special context of classical probability with equally probably cases some basic notions of probability theory

-   Today I want to establish theses notions a bit more generally and define them in mathematical terms independent of how probability is interpreted.

-   We formulate the concepts of probability theory in a way where probabilities are given and only have to fulfill certain axioms or rules

-   These rules will encompass what we discussed in the last lecture as one special case.

    -   Random Experiment When we use the term (random) experiment we mean a process leading to an uncertain outcome.

    -   Sample space The collection of all possible outcomes of an experiment is called the sample space

    -   Basic outcome, event, simple event When we talk of a basic outcome we mean the possible outcome of a random experiment. An event is a subset of basic outcomes. Any event which consists of a single outcome in the sample space is called a simple event.

    -   Probability Probability is a measure of how likely an event of an experiment is

-   Note that in probability theory probabilities are numbers between 0 and 1 which fulfill certain rules and are defined with respect to a random experiment and a sample space

-   From the viewpoint of theory they are considered as given

-   Probabilities in probability theory are of the same nature as distances in geometry

-   In probability theory we need not assume anything about how probabilities are measured

-   In actual applications of probabilities we often need sophisticated statistical methods to determine probabilities

-   Probability Given a (discrete) sample space S the probability assigned to events in this sample space must always fulfill three rules:

    -   P($S$) = 1, where $S$ is the sample space

    -   For any event A ∈ $S$, 0 ≤ P(A) ≤ 1. The probability of an event can never be negative or larger than 1

    -   The probability of the union of two events A and B is always smaller or equal to the sum of the probability of these events looked at in isolation: P(A ∪ B) ≤ P(A) + P(B)

-   Create a virtual coin:

    ```{r}
    coin <- c("H","T")
    ```

-   and a function to toss it:

    ```{r}
    toss_coin <- function() {coin <- c("H","T")
    sample(coin,size=1) }
    ```

    ```{r}
    tosses_100 <- replicate(100, toss_coin())

    library(ggplot2)
    qplot(tosses_100)
    ```

-   toss 1000 times

    ```{r}
    tosses_1000 <- replicate(1000,toss_coin())

    qplot(tosses_1000)
    ```

-   toss 1 million times

    ```{r}
    tosses_1mill <- replicate(10^6,toss_coin())

    qplot(tosses_1mill)
    ```

-   apple prices read:

    ```{r}
    aapl_prices <- read.csv("C:\\Users\\Rakete\\OneDrive\\Radna površina\\aapl_prices.csv")
    ```

-   Lets inspect the object using the function head() with the parameter value n = 5. This will show me the first 10 rows of the datafile.

    ```{r}
    head(aapl_prices,n=5)
    ```

-   An atomic vector is just a simple vector of data. For example our die constructed in the first lecture is an instance of this.

    ```{r}
    die <- 1:6
    ```

-   If you explicitly specify number as integers, you need to append the letter L to the number

    ```{r}
    die_int <- c(1L,2L,3L,4L,5L,6L)

    typeof(die_int)
    ```

    ```{r}
    list_example <- list(100:130, "R", list(TRUE, FALSE))
    ```

-   If you select two or more columns from a data frame, R will return a new data frame, like in

    ```{r}
    aapl_prices[1:2,1:6]
    ```
